music

### 2020-01-23 analysis/22_symphony.Rmd
more symphony foo

>>>>

```{r symphonymodules}
modules <- read_csv("../results/08_genes_modules.csv") %>%
  rename(modulecolor = `net$colors`)

candidategenes <- c("PRL", "PRLR", 
                 "VIP", "VIPR1", "VIPR2", 
                 "OXT", "AVP", "AVPR1A", "AVPR1B", 
                 "GNRH1","GNRHR", "NPVF",
                 "NR3C1", "NR3C2",
                 "ESR1", "ESR2", "AR",
                 "DIO2","LEPR", "DIO3", "DIO1","CYP19A1",
                 "HSPA14", "HSPA12A",
                 "PTGES3", "HSD11B2","DRD5", "DRD1", "DRD2","PGE1", "PGF")

modulescandidates <- modules %>% filter(gene %in% candidategenes)  %>%
  group_by(modulecolor)  %>%
  summarize(gene = str_c(gene, collapse = ", "))

modulescandidates

datapath <- "../results/"   # path to the data
datafiles <- dir(datapath, pattern = "*allvsd.csv") # get file names
datapathfile <- paste0(datapath, datafiles)

df <- datapathfile %>%
  setNames(nm = .) %>%
  map_df(~read_csv(.x, col_types = cols(), col_names = T), .id = "filename") %>% 
  mutate(tissue = sapply(strsplit(as.character(filename),'../results/06_'), "[", 2)) %>% 
  mutate(tissue = sapply(strsplit(as.character(tissue),'allvsd.csv'), "[", 1))  %>% 
  select(tissue, X1, everything())

df2 <-  df  %>%
  filter(X1 %in% c(candidategenes)) %>%
  pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "sample", values_to = "vsd") %>%
   mutate(sex = sapply(strsplit(as.character(sample),'_'), "[", 2)) %>%
   mutate(treatment = sapply(strsplit(as.character(sample),'_'), "[", 4))  %>%
   mutate(treatment = sapply(strsplit(as.character(treatment),'.NYNO'), "[", 1)) %>%
  mutate(bird = sapply(strsplit(as.character(sample),'_'), "[", 1)) %>%
  filter(treatment %in% charlevels) %>%
  select(bird, sex, treatment, tissue, X1, vsd) %>%
  mutate(tissue = fct_recode(tissue, "hypothalamus" = "hyp",
                    "pituitary" = "pit",
                    "gonads" = "gon"
                    )) %>%
  rename(gene = X1) %>%
  drop_na() %>%
  droplevels()
head(df2)

df2$treatment <- factor(df2$treatment, levels = alllevels)
df2$tissue <- factor(df2$tissue, levels = tissuelevels)
df2$gene <- factor(df2$gene)

df3 <- df2 %>% 
  mutate(treatment = fct_relevel(treatment, charlevels)) %>% 
  group_by(treatment, tissue, gene)  %>% 
  summarize(m = mean(vsd, na.rm = T), se = sd(vsd,  na.rm = T)/sqrt(length(vsd))) %>%
  mutate(image = "../figures/images/DoveParentsRNAseq_note.png") 
head(df3)

d4 <- left_join(df3, modules, by = "gene")

for (i in levels(d4$tissue)) {
  p <-  d4 %>%
    filter(tissue == i) %>%
    ggplot(aes(x = treatment, y = m)) +
    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
    geom_point(size = 1, aes(color = gene)) +
    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
    scale_alpha_manual(values = c(0.5, 1)) +
    labs(subtitle = i, y = "average expression", x = "parental stage") +
    facet_wrap(~modulecolor, nrow = 2) +
    theme_B3() +
    theme(legend.position = "bottom")
 print(p)
}

d4 %>%
  filter(modulecolor != "turquoise",
         tissue == "pituitary") %>%
  droplevels() %>% 
    ggplot(aes(x = treatment, y = m)) +
    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
    geom_point(size = 1, aes(color = gene)) +
    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
    scale_alpha_manual(values = c(0.5, 1)) +
    labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "average expression", x = "parental stage") +
    facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
    theme_B3() +
    theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
    guides(color = guide_legend(nrow = 2, byrow = T))

d4 %>%
  filter( modulecolor != "turquoise",
          # modulecolor != "greenyellow",
          tissue == "pituitary") %>%
  droplevels() %>% 
    ggplot(aes(x = treatment, y = m)) +
  geom_image(aes(image=image), size = 0.15) +
  facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
  labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "gene expression", x = "parental stage") +
    facet_wrap(~gene, scales = "free_y", nrow = ) +
    theme_B3() +
    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

d4 %>%
  filter( modulecolor != "turquoise",
          # modulecolor != "greenyellow",
          tissue == "pituitary") %>%
  droplevels() %>% 
    ggplot(aes(x = treatment, y = m)) +
  geom_image(aes(image=image), size = 0.1) +
  labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "gene expression", x = "parental stage") +
    theme_B3() +
    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_log10()

## outline

- data-driven
- hypothesis-drive
- sonically-driven
- visually-driven
- statistcally-driven

Data and hypothesis driven research for the visually, sonically, and statistcially inclined biologists
